---
title: "Eating disorder paper inference statistics"
author: "SP"
date: "`r Sys.Date()`"
output: 
  workflowr::wflow_html: 
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    df_print: paged
    fig_width: 11
    fig_height: 8
    fig_caption: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = TRUE)
# knitr::opts_knit$set(root.dir = dirname(getwd()))
```

```{r libs, include=FALSE, echo=FALSE}
do_source_data_creation <- FALSE
source("code/eating_disorder_data_proc.R")
load_inst_pkgs("tidyverse", "psych", "lmtest", "janitor", "knitr")
```

Note: Reported here are the uncorrected p-values. Controlling for FDR is done within the adjoined tables via Benjamini–Hochberg procedure.

# Pre-pandemic vs Covid pandemic

## Eating Disorders {.tabset .tabset-pills}

### Eating Disorder (ED) cases
```{r}
kruskal.test(icd_f50 ~ covid_pan, df_ed)
```
### ED-subgroup Anorexia (AN)
```{r}
kruskal.test(f50_type == "Anorexie" ~ covid_pan, df_ed)
```
### ED-subgroup Bulimia (BN)
```{r}
kruskal.test(f50_type == "Bulimie" ~ covid_pan, df_ed)
```
### ED-subgroup other EDs (oED)
```{r}
kruskal.test(f50_type == "Essstörung Sonst." ~ covid_pan, df_ed)
```

## clinical presentation {.tabset .tabset-pills}

### age
#### all cases
```{r}
oneway.test(age_adm ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
oneway.test(age_adm ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
oneway.test(age_adm ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### sex
#### all cases
```{r}
kruskal.test(sex ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
kruskal.test(sex ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
kruskal.test(sex ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

## comorbidities {.tabset .tabset-pills}

### total comorbidities / number of diagnoses
#### all cases
```{r}
oneway.test(n_icd_code ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
oneway.test(n_icd_code ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
oneway.test(n_icd_code ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### psychiatric comorbidities (co-occuring F-diagnoses)
#### all cases
```{r}
oneway.test(n_icd_f ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
oneway.test(n_icd_f ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
oneway.test(n_icd_f ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### somatic comorbidities (co-occuring non-F-diagnoses)
#### all cases
```{r}
oneway.test(n_icd_other ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
oneway.test(n_icd_other ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
oneway.test(n_icd_other ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### Depressive disorders (DD)
#### all cases
```{r}
kruskal.test(icd_dd ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
kruskal.test(icd_dd ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
kruskal.test(icd_dd ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### Anxiety disorders (AD)
#### all cases
```{r}
kruskal.test(icd_ad ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
kruskal.test(icd_ad ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
kruskal.test(icd_ad ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

### combined ED+DD+AD
```{r}
kruskal.test(icd_f50_dd_ad ~ covid_pan, df_ed)
```

## healthcare utilization {.tabset .tabset-pills}

### Length of stay
#### all cases
```{r}
oneway.test(length_stay_brutto ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
oneway.test(length_stay_brutto ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
oneway.test(length_stay_brutto ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```
#### cases with combined ED+DD+AD
```{r}
oneway.test(length_stay_brutto ~ covid_pan, df_ed %>% filter(icd_f50_dd_ad))
```

### re-admissions (v1)
```{r echo=FALSE}
re_adm_table <- 
  bind_rows(baseline = df_ed_re_baseline %>% tabyl(is_first_case, icd_f50),
            dur_covid = df_ed_re_covid %>% tabyl(is_first_case, icd_f50),
            .id = "period") %>% 
  rename(F50minus = "F50-", F50plus = "F50+") %>% 
  mutate(combined = F50minus + F50plus) %>% 
  arrange(is_first_case)
```
#### all cases
```{r}
chisq.test(matrix(re_adm_table$combined, nrow = 2))
```
#### ED cases only
```{r}
chisq.test(matrix(re_adm_table$F50plus, nrow = 2))
```
#### non-ED cases only
```{r}
chisq.test(matrix(re_adm_table$F50minus, nrow = 2))
```

### early re-admissions (within 6 months of last discharge)
#### all cases
```{r}
kruskal.test(re_adm_soon_nona ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
kruskal.test(re_adm_soon_nona ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
kruskal.test(re_adm_soon_nona ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```

## discharge medication {.tabset .tabset-pills}

### psychotropic medication
#### all cases
```{r}
kruskal.test(psych_med ~ covid_pan, df_ed)
```
#### ED cases only
```{r}
kruskal.test(psych_med ~ covid_pan, df_ed %>% filter(icd_f50 == "F50+"))
```
#### non-ED cases only
```{r}
kruskal.test(psych_med ~ covid_pan, df_ed %>% filter(icd_f50 == "F50-"))
```


# Within Covid pandemic

## School closure severity
### all cases
```{r}
kruskal.test(icd_f50 ~ covid_lockd_sev, df_ed_covid)
```
### AN cases only
```{r}
kruskal.test(f50_type == "Anorexie" ~ covid_lockd_sev, df_ed_covid)
```
### BN cases only
```{r}
kruskal.test(f50_type == "Bulimie" ~ covid_lockd_sev, df_ed_covid)
```
### oED cases only
```{r}
kruskal.test(f50_type == "Essstörung Sonst." ~ covid_lockd_sev, df_ed_covid)
```
### cases with combined ED+DD+AD
```{r}
kruskal.test(icd_f50_dd_ad ~ covid_lockd_sev, df_ed_covid)
```

## Covid pandemic phases {.tabset .tabset-pills}

### trend of ED share over periods
#### all cases
```{r}
cor.test(df_ed_covid$icd_f50 %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")
```
#### AN cases only
```{r}
cor.test((df_ed_covid$f50_type == "Anorexie") %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")
```
#### BN cases only
```{r}
cor.test((df_ed_covid$f50_type == "Bulimie") %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")
```
#### oED cases only
```{r}
cor.test((df_ed_covid$f50_type == "Essstörung Sonst.") %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")
```
#### cases with combined ED+DD+AD
```{r}
cor.test(df_ed_covid$icd_f50_dd_ad %>% as.numeric(),
         df_ed_covid$covid_period_i %>% as.numeric(),
         method = "spearman")
```

### comparison of pandemic phase with same months of the respective baseline period
```{r echo=FALSE}
create_covid_period_baseline_df <- function(i = 1L) {
  df_ed %>% 
    # baseline period
    filter(covid_pan == "pre_covid", 
           as_date(adm_date_year0) %in% (df_covid_periods_is_lockd %>% 
                                           filter(period_i_is_lockd == i) %>% 
                                           use_series(period_days) %>% 
                                           extract2(1))) %>% 
    # covid_period
    bind_rows(df_ed %>% filter(as.numeric(covid_period_i) == i+1))
}
```
#### phase 1: 1st open period
```{r}
kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(1))
```
#### phase 2: 1st school closure period
```{r}
kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(2))
```
#### phase 3: 2nd open period
```{r}
kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(3))
```
#### phase 4: 2nd school closure period
```{r}
kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(4))
```
#### phase 5: 3rd open period
```{r}
kruskal.test(icd_f50 ~ covid_pan, create_covid_period_baseline_df(5))
```

### Granger causality model: AN by school closure severity
```{r}
grangertest(AN_prop ~ as.numeric(covid_lockd_sev), order = 6, data = df_ed_covid_monthly)
```
#### Impulse response function
```{r echo=FALSE}
# test for stationarity
library(tseries)
adf.test(df_ed_covid_monthly$AN_prop, k = 6)

# IRF
library(vars)
ed_covid_monthly_irf <- df_ed_covid_monthly %>% 
  transmute(AN_prop, sev = as.numeric(covid_lockd_sev)) %>% 
  as.ts() %>% 
  VAR(p = 6, type = "const") %>% 
  irf()

plot(ed_covid_monthly_irf)
```
